### 2024-07-17 publications.tex
Update

\section{Publications}

\subsection{Journal Publications -- \href{https://ui.adsabs.harvard.edu/search/filter_database_fq_database=OR&filter_database_fq_database=database\%3A\%22astronomy\%22&fq=\%7B!type\%3Daqp\%20v\%3D\%24fq_database\%7D&fq_database=(database\%3A\%22astronomy\%22)&q=\%3Dauthor\%3A(\%22Kim\%2C\%20Jeong-Gyu\%22)&sort=date\%20desc\%2C\%20bibcode\%20desc}{\color{dodgerblue}{ADS
Search}}}

% ** Important publications ;

\# Led by a student under direct supervision

\begin{benumerate}{22}
\item \textit{Metallicity Dependence of Pressure-Regulated Feedback-Modulated Star Formation in the TIGRESS-NCR Simulation Suite} \\ Kim, C.-G., Ostriker, E. C., \textbf{Kim, J.-G.}, Gong, M., \& LtU collaboration \textit{2024, ApJ accepted}
\item \textit{Geometry, Dissipation, Cooling, and the Dynamical Evolution of Wind-Blown Bubbles} \\ Lancaster, L., Ostriker, E. C., Kim, C.-G., \textbf{Kim, J.-G.}, \& Bryan, G. \textit{2024, ApJ, 970, 18}
\item \textit{Implementation of Chemistry in the Athena++ Code} \\ Gong, M. et al. (incl.
\textbf{JGK}), 2023, ApJS, 268, 42
\item \textit{Physical Modeling of Dust Polarization from Magnetically Enhanced Radiative
    Torque (MRAT) Alignment in Protostellar Cores with POLARIS} \\ Giang, N. C., Hoang,
  T., \textbf{Kim, J.-G.}, \& Tram, L. N. 2023, MNRAS, 520, 3788
\item \textit{Introducing TIGRESS-NCR: I. Co-regulation of Multiphase ISM and
Star Formation Rates} \\ Kim, C.-G., \textbf{Kim, J.-G.}, Gong, M., \& Ostriker,
E. C. 2023, ApJ, 946, 3
\item \textit{Photochemistry and Heating/Cooling of the Multiphase Interstellar
Medium with UV Radiative Transfer in Magnetohydrodynamic Simulations} \\
\textbf{Kim, J.-G.}, Gong, M., Kim, C.-G., \& Ostriker, E. C. 2023, ApJS, 264, 10
\item \textit{Slow Star Formation in the Milky Way: Theory Meets Observations}\\
  Evans, Neal J., II, \textbf{Kim, J.-G.}, \& Ostriker, E. C. 2022, ApJL, 929L,
  18E
\item \textit{Star Formation Regulation and Self-Pollution by Stellar Wind
    Feedback}\\
  Lancaster, L., Ostriker, E. C., \textbf{Kim, J.-G.}, \& Kim, C.-G. 2021, ApJL,
  922L, 3L
\item \textit{Efficiently Cooled Stellar Wind Bubbles in Turbulent Clouds:
    II. Validation of Theory with Hydrodynamic Simulations}\\
  Lancaster, L., Ostriker, E. C., \textbf{Kim, J.-G.}, \& Kim, C.-G. 2021, ApJ,
  914, 90L
\item \textit{Efficiently Cooled Stellar Wind Bubbles in Turbulent Clouds:
I. Fractal Theory and Application to Star-Forming Clouds}\\
  Lancaster, L., Ostriker, E. C., \textbf{Kim, J.-G.}, \& Kim, C.-G., 2021, ApJ,
914, 89L
\item \textit{Star Formation Efficiency and Dispersal of Giant Molecular
    Clouds with UV Radiation Feedback: Dependence on Gravitational Boundedness
    and Magnetic Fields} \\ \textbf{Kim, J.-G.}, Ostriker, E. C., \& Filippova,
  N. 2021, ApJ, 911, 128K
\item \textit{The environmental dependence of the $X_{\mathrm CO}$ conversion
    factor}\\ Gong, M., Ostriker, E. C., Kim, C.-G., \& \textbf{Kim,
    J.-G.} 2020, ApJ, 903, 142
\item \textit{Factories of CO-dark gas: molecular clouds with limited star
formation efficiencies by far-ultraviolet feedback} \\
  Inoguchi, M., Hosokawa, T., Mineshige, S., \& \textbf{Kim, J.-G.} 2020, MNRAS,
  497, 5061I
\item \textit{\# Diffuse Ionized Gas in Simulations of Multiphase, Star-forming
    Galactic Disks}\\
  Kado-Fong, E., \textbf{Kim, J.-G.}, Ostriker, E. C., \& Kim, C.-G. 2020, ApJ,
  897, 143
\item \textit{Modeling UV Radiative Feedback from Massive Stars: III. Escape
    of Radiation from Star-Forming Giant Molecular Clouds} \\ \textbf{Kim,
    J.-G.}, Kim, W.-T., \& Ostriker, E. C. 2019, ApJ, 883, 102
\item \textit{Modeling UV Radiative Feedback from Massive Stars: II. Dispersal
of Star-Forming Giant Molecular Clouds by Photoionization and Radiation
Pressure} \\ \textbf{Kim, J.-G.}, Kim, W.-T., \& Ostriker, E. C. 2018, ApJ, 859,
68
\item \textit{Modeling UV Radiative Feedback from Massive Stars: I.
Implementation of Adaptive Ray Tracing Method and Tests}\\ \textbf{Kim, J.-G.},
Kim, W.-T., Ostriker, E. C.,\& Skinner A. M. 2017, ApJ, 851, 93
\item \textit{Disruption of Molecular Clouds by Expansion of Dusty H II
Regions}\\ \textbf{Kim, J.-G.}, Kim, W.-T., \& Ostriker, E. C. 2016, ApJ, 819,
137
\item \textit{Instability of Magnetized Ionization Fronts Surrounding H II
regions}\\ \textbf{Kim, J.-G.}, Kim, W.-T. 2014, ApJ, 797, 135
\item \textit{Nature of Wiggle Instability of Galactic Spiral Shocks}\\ Kim,
W.-T., Kim, Y., \& \textbf{Kim, J.-G.} 2014, ApJ, 789, 68
\item \textit{Instability of Evaporation Fronts in the Interstellar Medium}\\
  \textbf{Kim, J.-G.}, Kim, W.-T. 2013, ApJ, 779, 48
\item \textit{Gravitational Instability of Rotating, Pressure-Confined,
    Polytropic Gas Disks with Vertical Stratification} \\ \textbf{Kim, J.-G.},
  Kim, W.-T., Seo Y. M., \& Hong, S. S. 2012, ApJ, 761, 131
\end{benumerate}

\subsection{Papers in Preparation}
\begin{enumerate}
\item[] \textit{Structure and Dynamics of Density-Bounded H II Regions with Photoevaporative Outflows} \\ \textbf{Kim, J.-G.}, Ostriker, E. C., Lancaster, L., \& Kim, C.-G. \textit{2024, in preparation}
\item[] \textit{Ultraviolet Radiation Fields in Star-Forming Disk Galaxies: Numerical Simulations with TIGRESS-NCR} \\ Linzer, N., \textbf{Kim, J.-G.}, Kim, C.-G., \& Ostriker, E. C. \textit{2024, ApJ submitted}

% \item[] \textit{What Regulates Galactic Star Formation Rates? A View from New
%     TIGRESS Simulations} \\ Kim, C.-G., \textbf{Kim, J.-G.}, Gong, M., \& Ostriker,
%   E. C. \textit{2022, to be submitted}
% \item[] \textit{Destruction of Giant Molecular Clouds by Stellar Feedback: I.
%     Relative Importance of Radiation, Winds, and Supernovae} \\ \textbf{Kim,
%     J.-G.}, Lancaster, L., Kim, C.-G., \& Ostriker, E. C. \textit{2022, in
%     prep.}
% \item[] \textit{The Lyman Continuum and Far-UV Interstellar Radiation Field in
%     Simulations of Multiphase, Star-forming Galactic Disks} \\ Linzer, N.,
%   Ostriker, E. C., \textbf{Kim, J-.G.}, \& Kim, C.-G. \textit{2022, in prep.}
% \item[] \textit{The Effects of Magnetic Fields and H II regions on the Dynamics
% of Stellar Wind Bubbles} \\ Lancaster, L., Ostriker, E. C.,
%   \textbf{Kim, J.-G.}, \& Kim, C.-G. \textit{2022, in prep.}
% \item[] \textit{r-process Enrichment During the Formation of a Globular Cluster: Case
%     of M15} \\
%   Hotokezaka, K., \textbf{Kim, J.-G.}, Beniamini, P., \& Cen, R. \textit{in prep.}
 % \item[] \textit{Recovery of Cloud-Scale Star Formation Rate Based on Ionizing
 %    Luminosity: Calibration for Escape of Radiation} \\ \textbf{Kim, J.-G.} \&
 %  Ostriker, E. C. \textit{in prep.}
\end{enumerate}

%%% Local Variables: %%% mode: latex %%% TeX-master: t %%% End: